---
title: radolanr
output:
  github_document:
    number_sections: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
# set general behaviour of the chunks here.. you may still modify it later in the chunks
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  echo = TRUE, 
  results = "hide", # hide outputs of a the code (warnings ect. will still be printed; comand to unable in individual chunks: results="markup")
  error=FALSE, # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
  warning=FALSE,  # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
  message=FALSE  # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
)
```

# Description
This package helps to download series of radolan data and to create timeseries of this data for given centroids provided as shape files.
This package provides some usefull wrappers around the rdwd package to make more convinient processes. Outputs may likewise be used by rdwd package functions. 

# Usage

## Install Package
```{r, eval = FALSE}
# This package is not yet on CRAN. The easiest way to get sapflowr is to install it via github:
library(devtools)
install_github("Blizno/radolanr") # this will install the version on "main"
```

```{r, eval = FALSE}
# At the moment the package is a private package, you will need credentials
https://rdrr.io/cran/remotes/man/install_github.html recomends: 
# "
# To install from a private repo, use auth_token with a token
# from https://github.com/settings/tokens. You only need the
# repo scope. Best practice is to save your PAT (Personal Access Token) in env var called
# GITHUB_PAT.
# "
install_github("Blizno/radolanr", auth_token = "abcdefasdf")

# pottentially you may also use the https link if you are logged into your github account at the same time (e.g. via www.github.de in our browser)
```

## load the package
```{r, eval = TRUE}
library(radolanr)
```


## developing the package
Dependencies are not automatically writen to DESCRIPTION and thus will not be loaded when the package is installed via install_github(). Run the following code to automatically write the used libraries (documented as Roxygen '# @import ...) to DESCRIPTION what will automatically install them as dependencies.
```{r}
library(attachment)
library(desc)
library(devtools)

# Auto-detect dependencies and update DESCRIPTION
attachment::att_amend_desc()

# # Ensure dependencies install when package is installed
# devtools::install_deps(dependencies = TRUE)

message("✅ Dependencies updated and installed!")

# run check, This ensures there are no missing dependencies.
devtools::check()

```
