---
title: use case 2 - get rain distribution along the Elbe River
output:
  github_document:
    number_sections: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
# set general behaviour of the chunks here.. you may still modify it later in the chunks
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  echo = TRUE, 
  results = "hide", # hide outputs of a the code (warnings ect. will still be printed; comand to unable in individual chunks: results="markup")
  error=FALSE, # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
  warning=FALSE,  # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
  message=FALSE  # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
)
```

# Introduction

# Install packages
```{r}
#library(rdwd)
#rdwd::updateRdwd() # --> installes the last version, developement version on the github is used..

# library(rdwd)
# #install.packages('dwdradar')
# library(dwdradar)
library(radolanr)

library(ggplot2)
library(tidyverse) 
library(reshape2)
library(terra) # for the raster and vector data treatment
```

## DATENBESCHREIBUNG
--> refer to UseCase1

## Process data
Access the radolan data latest dataset. Plotting will be undergone according to the rdwd package
```{r}
radp <- radolanr::dataDWDPerDay(mode = "latest", addMetaData = TRUE)
```



# extract values from raster file and save to given polygons
extract values at specific point with raster.. 

Todo: bisher stimmen diese Werte nicht mit denen überein die ich in der App ausgebe --> also mit denen die von https:---recent erstellt werden. Die Daten auf dem opendata.dwd scheinen 1h aktueller als die Daten auf dem cdc-ftpserver

```{r}
transsect_rain <- radolanr::radolan2polygon(
  locations_polygon_path = "data_raw/transsect_weinbau_seusslitz_pillnitz_polygons_4326.shp",
  radolan_raster = radp, # radolan raster file
  saveCentroidsVector = FALSE
)


```


# Plots

plot metadata stored along the raster (time of access and acutality of the radolan data). They are added by radolanr() if radolanr::dataDWDPerDay(mode = "latest", addMetaData = TRUE)
```{r}
metags(radp)
```

Plot Radolan data
```{r}
#plotRadar(radp, main=paste("mm in 24 hours preceding"), project=FALSE)
```





# Plot results using Leaflet

## Changing legend based on the background, show values when overing over map
```{r}
library(leaflet)
library(sf)            # For spatial data
library(terra)         # For raster handling
library(RColorBrewer)  # For color palettes


# Define a color palette for rain amounts (based on polygons by default)
pal_rain <- colorNumeric(palette = "YlGnBu", domain = transsect_rain$rain_mm)

# Define a color palette for radp (raster) values
pal_radp <- colorNumeric(palette = "YlGnBu", domain = values(radp), na.color = "transparent")

# set timestamp
radolan_timestamp <- transsect_rain[1]$timestamp_radolan

# Create the Leaflet map
p1 <- leaflet() %>%
  # Base Layers: OpenStreetMap and RADP Raster
  addTiles(group = "OpenStreetMap") %>%  # Default basemap

  addRasterImage(radp, colors = pal_radp, opacity = 0.7, group = "RADP Raster") %>%  # Raster layer
  
  
  # Polygon Layer with Popups on Hover
  addPolygons(
    stroke = TRUE,               # Borderlines
    color = "black",             # Borderline color
    opacity = 1,                 # Borderline transparency
    data = transsect_rain,       # Polygon layer
    weight = 1,                  # Line thickness
    smoothFactor = 0.3,          # Smooth the polygon borders
    fillOpacity = 0.8,           
    fillColor = ~pal_rain(rain_mm),  # Color based on rain_mm
    label = ~paste0("Rain: ", rain_mm, " mm"),  # Show value on hover
    highlight = highlightOptions(weight = 2, color = "red", fillOpacity = 0.9) # Highlight effect on hover
  ) %>%


  # Add Layer Control to Toggle between OpenStreetMap and RADP Raster
  addLayersControl(
    baseGroups = c("OpenStreetMap", "RADP Raster"),  # User can switch backgrounds
    overlayGroups = c("Polygons"),  # Polygons always visible
    options = layersControlOptions(collapsed = FALSE)
  ) %>%

  # Add Dynamic Legend (changes based on background)
  addLegend(
    pal = pal_rain, values = transsect_rain$rain_mm, opacity = 0.8,
    labFormat = labelFormat(big.mark = "'"),
    title = paste0("Regen<br>Summe 24h<br>(mm)<br>RADOLAN<br>",radolan_timestamp),
    group = "OpenStreetMap"  # Legend applies when OpenStreetMap is active
  ) %>%
  addLegend(
    pal = pal_radp, values = values(radp), opacity = 0.8,
    labFormat = labelFormat(big.mark = "'"),
    title = "Regen<br>Summe 24h<br>(mm)<br>RADOLAN (Raster)",
    group = "RADP Raster"  # Legend applies when RADP Raster is active
  )

# Show the map
p1

```


# Export Data to be used e.g. in QGIS
```{r}
# terra::writeRaster(radp, "radolan4326.TIFF")
# terra::writeVector(transsect_rain, "transsect_weinbau_seusslitz_pillnitz_polygons_rain24h_4326.shp", overwrite = TRUE)
```

# Options: load shapefiles from a google-drive folder
```{r}



```

# Options: load shapefiles from QFieldCloud
```{r}



```
