---
title: use case 2 - get rain distribution along the Elbe River
output:
  github_document:
    number_sections: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
# set general behaviour of the chunks here.. you may still modify it later in the chunks
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  echo = TRUE, 
  results = "hide", # hide outputs of a the code (warnings ect. will still be printed; comand to unable in individual chunks: results="markup")
  error=FALSE, # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
  warning=FALSE,  # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
  message=FALSE  # results = 'hide' option doesn't prevent other messages to be printed. To hide them like this, they will still be printed to console
)
```

# Introduction

# Install packages
```{r}
#library(rdwd)
#rdwd::updateRdwd() # --> installes the last version, developement version on the github is used..
library(rdwd)

#install.packages('dwdradar')
library(dwdradar)
library(radolanr)

library(ggplot2)

library(tidyverse) 

library(reshape2)

library(terra) # for the raster and vector data treatment
```

# load vector file to extract data
```{r}
# read shapefiles of target areas
my.centroids <- terra::vect("data_raw/transsect_weinbau_seusslitz_pillnitz_4326.shp")
```


## DATENBESCHREIBUNG
--> refer to UseCase1

## Process data
Access the radolan data latest dataset. Plotting will be undergone according to the rdwd package
```{r}
radp <- radolanr::dataDWDPerDay(mode = "latest", addMetaData = TRUE)
```



# extract values from raster file
extract values at specific point with raster.. 

Todo: 
- bisher stimmen diese Werte nicht mit denen überein die ich in der App ausgebe --> also mit denen die von https:---recent erstellt werden. Die Daten auf dem opendata.dwd scheinen 1h aktueller als die Daten auf dem cdc-ftpserver

# start with shape files, calculate centroids, extract calculate back
# load vector file to extract data

I will use terra alongside with geos package. Use terra if working with raster & vector data (most modern for raster). Use sf if doing vector operations (widely used, but heavier dependencies). Use geos for fast geometry computations (complements terra well).

```{r}
radolan2polygon <- function(locations_polygon_path, radolan_raster, use_metadata = TRUE, silent = TRUE, saveCentroidsVector = FALSE, centroidsVectorPath = ""){
  # read shapefiles of target areas
  shp <- terra::vect(locations_polygon_path)
  my.centroids <- terra::centroids(shp)

  # # Optional/Todo:
  # # Convert SpatVector to geos_geometry and compute centroids that are guaranteed to be inside polygons (critical for irregular    polygons). This is not guaranteed for terra:centroides. However geos not natevly works with terra object so conversions are necessary.
  # my.centroids.inside <- geos::geos_point_on_surface(shp)

  if (saveCentroidsVector == TRUE){
    terra::writeVector(my.centroids, "transsect_weinbau_seusslitz_pillnitz_polygons_centroids_4326.shp", overwrite = TRUE)
  }

  # Extract raster values to list object
  r.vals <- terra::extract(radp, my.centroids, na.rm = FALSE)  # !! changed raster extract to terra:extract to be consistent here -->     maybe change this in UseCase1 too ; I think na should not be removed as this will change the ranking and number of extracted     values, but na.rm seems not to have an influence in the function, i.e. if polygons are outside the radolan map there will be NA in the resulting table, no matter if na.rm was TRUE or FALSE
  if (silent == FALSE){
    print(r.vals)
  }


  # Extract function returns a data frame with an ID column, so merge properly
  my.centroids$rain_mm <- r.vals[, 2]  # Second column contains raster values

  # Add metadata
  my.metadata       <- terra::metags(radp) # a names character vector
  timestamp_radolan <- my.metadata["timestamp_radolan"] # will give timestamp of radolan data in ISO8601 Format: "2025-03-26   08:50:00+0100" as character string (added by radolanr::dataDWDPerDay(mode = "latest", addMetaData = TRUE))


  # create output
  shp$rain_mm <- r.vals[,2]
  shp$timestamp_radolan <- timestamp_radolan
  
  return(shp)
  
}


transsect_rain <- radolan2polygon(
  locations_polygon_path = "data_raw/transsect_weinbau_seusslitz_pillnitz_polygons_4326.shp",
  radolan_raster = radp, # radolan raster file
  saveCentroidsVector = FALSE
)

```


# Plots

plot metadata stored along the raster (time of access and acutality of the radolan data). They are added by radolanr() if radolanr::dataDWDPerDay(mode = "latest", addMetaData = TRUE)
```{r}
metags(radp)
```

Plot Radolan data
```{r}
plotRadar(radp, main=paste("mm in 24 hours preceding"), project=FALSE)
```





# Plot results

# Plot results using Leaflet

## Changing legend based on the background, show values when overing over map
```{r}
library(leaflet)
library(sf)            # For spatial data
library(terra)         # For raster handling
library(RColorBrewer)  # For color palettes

# Load the polygon shapefile
# (Replace this with your actual spatial data)
# transsect_rain <- st_read("path_to_shapefile.shp")

# Load the raster layer
# (Replace this with your actual raster data)
# radp <- rast("path_to_raster.tif")

# Define a color palette for rain amounts (based on polygons by default)
pal_rain <- colorNumeric(palette = "YlGnBu", domain = transsect_rain$rain_mm)

# Define a color palette for radp (raster) values
pal_radp <- colorNumeric(palette = "YlGnBu", domain = values(radp), na.color = "transparent")

# Create the Leaflet map
p1 <- leaflet() %>%
  # Base Layers: OpenStreetMap and RADP Raster
  addTiles(group = "OpenStreetMap") %>%  # Default basemap

  addRasterImage(radp, colors = pal_radp, opacity = 0.7, group = "RADP Raster") %>%  # Raster layer
  
  
  # Polygon Layer with Popups on Hover
  addPolygons(
    stroke = TRUE,               # Borderlines
    color = "black",             # Borderline color
    opacity = 1,                 # Borderline transparency
    data = transsect_rain,       # Polygon layer
    weight = 1,                  # Line thickness
    smoothFactor = 0.3,          # Smooth the polygon borders
    fillOpacity = 0.8,           
    fillColor = ~pal_rain(rain_mm),  # Color based on rain_mm
    label = ~paste0("Rain: ", rain_mm, " mm"),  # Show value on hover
    highlight = highlightOptions(weight = 2, color = "red", fillOpacity = 0.9) # Highlight effect on hover
  ) %>%


  # Add Layer Control to Toggle between OpenStreetMap and RADP Raster
  addLayersControl(
    baseGroups = c("OpenStreetMap", "RADP Raster"),  # User can switch backgrounds
    overlayGroups = c("Polygons"),  # Polygons always visible
    options = layersControlOptions(collapsed = FALSE)
  ) %>%

  # Add Dynamic Legend (changes based on background)
  addLegend(
    pal = pal_rain, values = transsect_rain$rain_mm, opacity = 0.8,
    labFormat = labelFormat(big.mark = "'"),
    title = "Regen<br>Summe 24h<br>(mm)<br>RADOLAN",
    group = "OpenStreetMap"  # Legend applies when OpenStreetMap is active
  ) %>%
  addLegend(
    pal = pal_radp, values = values(radp), opacity = 0.8,
    labFormat = labelFormat(big.mark = "'"),
    title = "Regen<br>Summe 24h<br>(mm)<br>RADOLAN (Raster)",
    group = "RADP Raster"  # Legend applies when RADP Raster is active
  )

# Show the map
p1

```


# Export Data to be used e.g. in QGIS
```{r}
# terra::writeRaster(radp, "radolan4326.TIFF")
# terra::writeVector(transsect_rain, "transsect_weinbau_seusslitz_pillnitz_polygons_rain24h_4326.shp", overwrite = TRUE)
```

# Options: load shapefiles from a google-drive folder
```{r}



```

